Prediction of PIK3CA mutation with gene expression

### 2020-05-07 plosOne/plosOne.Rmd
introduction machine learning

Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype. @andre_2019_alpelisib_nengljmeda Prediction of PIK3CA mutation was done by gene expression data of TCGA.

- Predicting mutation by gene expression data
  * What for?  
  * Feasible?  
  * Cost?

- Varing frquency of PIK3CA mutation across cancer types  
  * High in endometrial breast  
  * not common other cancer type

Universial prediction using gene expression data across cancer type for certain kind of mutation. It's clinically useless now, but we want to explore the possibilities of the PIK3CA mutation prediction.

- RNAseq can be widely used. The mutation status directly prediction

- Previous study @way_2018_machine_cellreports

- Prediction of aberant activation of a certain pathway vs a specific mutation

- The proposed solution

@@

10845 cases were available both PIK3CA mutation and mRNAD expression data. 5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at modelinf process method. 33 cancer type dummy variables were included in predictor variables.

@@

## Dataset summary 
- PIK3CA prevalence
-

>>>>

Targeted thrapy becomes standard treatment in many cancer patients. To treat with the targeted therapy, the The patients should be confirmed the indicated genomic alterations in their cancer. Many direct test for the genomic alteration has been developed and prooved for their clical utility to discriminate which patients will be response the targeted therapy.

The machine learning approach has been actively researched recently to detect the genomic alterations. The machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image or gene expression data @wayMachineLearningDetects2018. The machine learning approach has limitations to apply to clinical practice. However the machine learning prediction can be a second best when the direct test fails.

PIK3CA is encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase which phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to make phosphatidylinositol 3,4,5-triphosphate (PIP~3~). Phosphatase and tensin homolog (PTEN) changes PIP~2~ to PIP~3~ in contrast PI3K. PIP~3~ is a second mesenger to activate protein kinase B (AKT) which is a serine/threonine-specific protein kinase. AKT inhibits apoptosis and promote cell proliferation.

Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype. @andre_2019_alpelisib_nengljmeda The PIK3CA mutation is second most driver mutation after TP53. The PIK3CA mutation is most frequantly founded in endometrial carcinoma (45%), and followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma and endocervical adenocarcinoma (20%) and colon adenocarcinoma (16%).

@@

10845 cases were available both PIK3CA mutation and mRNA expression data. 5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at modelinf process method. 33 cancer type dummy variables were included in predictor variables.

@@

## Dataset summary 
Prevalcence of PIK3CA was 0.11 in all cases. The PIK3CA prevalance of each cancer type was vary. The median prevalence of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).

### 2020-05-07 plosOne/plosOne.tex
introduction machine learning

Breast cancer with PIK3CA mutation has been approved to use PIK3CA
inhibitor in hormone receptor positive HER2 negative subtype. {[}1{]}
Prediction of PIK3CA mutation was done by gene expression data of TCGA.

@@

\begin{itemize}
\item
  PIK3CA prevalence
\item
\end{itemize}

>>>>

Targeted thrapy becomes standard treatment in many cancer patients. To
treat with the targeted therapy, the The patients should be confirmed
the indicated genomic alterations in their cancer. Many direct test for
the genomic alteration has been developed and prooved for their clical
utility to discriminate which patients will be response the targeted
therapy.

The machine learning approach has been actively researched recently to
detect the genomic alterations. Breast cancer with PIK3CA mutation has
been approved to use PIK3CA inhibitor in hormone receptor positive HER2
negative subtype. {[}1{]}

\begin{itemize}
\tightlist
\item
  PIK3CA summary
\end{itemize}

PIK3CA is encodes the p110\(\alpha\) catalytic subunit of
phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase which
phosphorylates phosphatidylinositol 4,5-biphosphate
(PIP\textsubscript{2}) to make phosphatidylinositol 3,4,5-triphosphate
(PIP\_\{3\}). Phosphatase and tensin homolog (PTEN) changes PIP\_\{2\}
to PIP\_\{3\} in contrast PI3K. PIP\_\{3\} is a second mesenger to
activate protein kinase B (AKT) which is a serine/threonine-specific
protein kinase. AKT inhibits apoptosis and promote cell proliferation.

The PIK3CA mutation is second most driver mutation after TP53. the
PIK3CA mutation is most frequantly founded in endometrial carcinoma
(45\%), and followed by breast invasive carcinoma (24\%), cervical
squamous cell carcinoma and endocervical adenocarcinoma (20\%) and colon
adenocarcinoma (16\%).

@@

Prevalcence of PIK3CA was 0.11 in all cases. The PIK3CA prevalance of
each cancer type was vary. The median prevalence of PIK3CA of each
cancer types was 0.03 (range 0-0.33) (Figure 1).
